Future challenges for Research Software Engineering

### 2025-09-24 future_challenges_for_RSEng.tex
split long sentence

Besides the direct impact in research software and further research output,
we see a number of risks.
The workshop participants suggested that the emergence of AI and automation might reduce critical thinking and imagination.
We address this in \autoref{sec:complexity}.
Another risk is that the widespread use of AI tools could make community resources
such as \eg\ Stack Overflow and Wikipedia (which are used to train AI models) poorer
due to fewer incentives to maintain them, although a recent survey suggests
these communities experience less activity from new users and an increased
average complexity of the questions posted~\cite{Burtch2024}.
The use of AI assistants in these communities could also compromise LLMs via `model collapse'
when AI-generated content becomes part of the training data~\cite{Shumailov2024}.

>>>>

Besides the direct impact in research software and further research output,
we see a number of risks.
The workshop participants suggested that the emergence of AI and automation might reduce critical thinking and imagination.
We address this in \autoref{sec:complexity}.
Another risk is that the widespread use of AI tools could make community resources
such as \eg\ Stack Overflow and Wikipedia (which are used to train AI models) poorer
due to fewer incentives to maintain them.
However, a recent survey suggests these communities experience less activity from new
users and an increased average complexity of the questions posted~\cite{Burtch2024}.
The use of AI assistants in these communities could also compromise LLMs via `model collapse'
when AI-generated content becomes part of the training data~\cite{Shumailov2024}.
